---
title: "Flavors of Cacao Analysis"
author: "Leopoldine Mirtil"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Source

The data used in this analysis is from the "Flavors of Cacao" database, made publicly available by Brady Brelinski, [link here](https://flavorsofcacao.com/chocolate_database.html) and was last updated March 3, 2024 at time of analysis. It  contains the ratings of nearly 2,700 dark chocolate bars. 


### Tasks

1. Determine the total amount of chocolate bars reviewed each year 
2. Identify the overall highest and lowest-rated chocolate bars
3. Determine the total ingredient count
4. Determine the companies and bean origin countries of the highest rated bars
5. Identify the characteristics of the most recent years' highest 4-Rated chocolate bars 


### Step 1: Import Data

#### Load Library

```{r, message=FALSE, echo=FALSE, warning=FALSE}
#for data manipulation
library(dplyr)
library(lubridate)
library(tidyr)
library(tidyverse)
library(readxl)

#for visualizations
library(ggplot2)
library(RColorBrewer)
library(wordcloud2)

#for documentation/markdown
library(knitr)
```


#### Import Data

```{r, message=FALSE, echo=FALSE}
cacao_df <- read_xlsx("Flavors of Cacao.xlsx", col_names=TRUE)
```


### Step 2 - Clean & Inspect Data

#### Check for Missing/NA Values

```{r}
sapply(cacao_df, function(x) sum(is.na(x)))
```


#### Check for Empty Values

```{r}
sapply(cacao_df, function(x) sum(x==""))
```


#### Remove Data

```{r}
#remove rows w/ missing/NA values
cacao_df <- drop_na(cacao_df)

#confirm removal
sapply(cacao_df, function(x) sum(is.na(x)))
```


#### Modify Ingredient Column

```{r}
#split ingredient columns: number & letters
cacao_df <- cacao_df %>% 
        separate(Ingredients, c('Ingredient Nums', 'Ingredients'), "-") 

#change data type of ingr nums
cacao_df$'Ingredient Nums' <- as.numeric(cacao_df$'Ingredient Nums')

#trim white space from ingredient (letters)
cacao_df$Ingredients <- str_trim(cacao_df$Ingredients)
 
#create dict for ingredients 
ingrd_dict <-  c('B'='Beans', 'S'='Sugar', 'S*'='Sweetener other than white cane or beet sugar', 
        'C'='Cocoa Butter', 'V'='Vanilla', 'L'='Lecithin','Sa'='Salt')


#replace ingredients letters using dictionary
cacao_df$Ingredients <- sapply(strsplit(cacao_df$Ingredients, ',\\s?'), function(x){paste(unlist(as.list(ingrd_dict)[x]), collapse = ',')})
```


#### View Modified Dataframe

```{r}
str(cacao_df)
```


```{r, echo=FALSE}
#export file
write.csv(cacao_df, 'cacao_df.csv', row.names = FALSE)
```


### Step 3: Analysis

#### Descriptive Analysis

```{r}
summary(cacao_df)
```

#### Unique Value Count of All Columns

```{r}
cacao_df %>%
  select(`Company (Manufacturer)`, `Review Date`, `Country of Bean Origin`, `Specific Bean Origin or Bar Name`, `Cocoa Percent`) %>%
  sapply(function(x) length(unique(x)))
```


#### Total Chocolate Bars Count by Rating

```{r}
cacao_df %>%
  select(Rating) %>%
  group_by(Rating) %>%
  summarise(count=n()) %>%
  arrange(desc(Rating))
```

#### Total Chocolate Bars Count by Year

```{r}
cacao_df %>%
  select(`Review Date`) %>%
  group_by(`Review Date`) %>%
  summarise(totalCount=n()) %>%
  arrange(desc(`Review Date`))
```

#### Total Ingredient Count

```{r}
cacao_df %>%
  select(Ingredients) %>%
  group_by(Ingredients) %>%
  separate_longer_delim(c(Ingredients), delim=",") %>%
  summarise(Total=n()) %>%
  arrange(desc(Total))
```


#### Highest Rated Chocolate Bar(s)

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, Rating) %>%
  filter(Rating == max(Rating)) %>%
  group_by(`Specific Bean Origin or Bar Name`) %>%
  arrange(`Specific Bean Origin or Bar Name`) %>%
  slice_head(n=1) 
```

#### Lowest Rated Chocolate Bar(s)

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, Rating) %>%
  filter(Rating == min(Rating)) 
```

#### Bean Origin Countries of 4-Rated Bars

```{r}
cacao_df %>%
  select(`Country of Bean Origin`, Rating) %>%
  filter(Rating == max(Rating)) %>%
  group_by(`Country of Bean Origin`) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
```


#### Companies of Highest-Rated Chocolate Bar(s)

```{r, message=FALSE}
cacao_df %>%
  select(`Company (Manufacturer)`, Rating) %>%
  filter(Rating == max(Rating)) %>%
  group_by(`Company (Manufacturer)`) %>%
  slice_head(n = 1) %>%
  arrange(`Company (Manufacturer)`)
```


#### Most Recent Years Analysis

##### 2023 Highest Rated Chocolate Bar(s)

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, `Review Date`, Rating, `Company (Manufacturer)`, `Country of Bean Origin`) %>%
  filter((`Review Date`==max(`Review Date`)) & (Rating==max(Rating)))
```


##### 2023 Lowest Rated Chocolate Bar(s)

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, `Review Date`, Rating, `Company (Manufacturer)`, `Country of Bean Origin`) %>%
  filter(`Review Date`==max(`Review Date`)) %>%
  filter(Rating==min(Rating))
```


##### Characteristics of 2023 Highest-Rated Bars

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`, Rating) %>%
  filter(Rating== max(Rating)) %>%
  group_by(`Specific Bean Origin or Bar Name`) %>%
  arrange(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`) %>%
  group_by(`Most Memorable Characteristics`) %>%
  separate_longer_delim(c(`Most Memorable Characteristics`), delim=",") %>%
  select(`Most Memorable Characteristics`) %>%
  summarize(count=n()) %>%
  arrange(desc(count)) 
```


##### Characteristics of 2023 Lowest-Rated Bars

```{r}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`, Rating) %>%
  filter(Rating == min(Rating)) %>%
  group_by(`Specific Bean Origin or Bar Name`) %>%
  arrange(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`) %>%
  group_by(`Most Memorable Characteristics`) %>%
  separate_longer_delim(c(`Most Memorable Characteristics`), delim=",") %>%
  select(`Most Memorable Characteristics`) %>%
  summarize(count=n()) %>%
  arrange(desc(count))
```


### Step 4: Visualization

```{r, echo=FALSE, warning=FALSE}
cacao_df %>%
  select(Rating) %>%
  group_by(Rating) %>%
  summarise(TotalCount=n()) %>%
    ggplot(aes(x=Rating, y=TotalCount))+
      geom_col(fill="brown4") +
      theme(axis.title.x=element_blank()) +
      ylab("Count") +
      ggtitle("Total Count by Rating")
```


```{r, echo=FALSE, warning=FALSE}
cacao_df %>%
  select(Ingredients) %>%
  group_by(Ingredients) %>%
  separate_longer_delim(c(Ingredients), delim=",") %>%
  summarise(TotalCount=n()) %>%
  arrange(desc(TotalCount)) %>%
  mutate(percentage=round((TotalCount*100/sum(TotalCount)), 2)) %>% 
  mutate(y_pos = cumsum(percentage)-0.5*percentage) %>%
    ggplot(aes(x="", percentage, fill=Ingredients))+
    ggtitle("Percentage of Ingredients Total") +
    geom_bar(width = 1, stat = "identity", color= "white", alpha=.5)+
    coord_polar("y", start=0)+
    scale_fill_manual(values=rainbow(9))+ 
    theme_void() 
```


```{r, echo=FALSE, warning=FALSE}
cacao_df %>%
  select(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`, Rating) %>%
  filter(Rating== max(Rating)) %>%
  group_by(`Specific Bean Origin or Bar Name`) %>%
  arrange(`Specific Bean Origin or Bar Name`, `Most Memorable Characteristics`) %>%
  group_by(`Most Memorable Characteristics`) %>%
  separate_longer_delim(c(`Most Memorable Characteristics`), delim=",") %>%
  select(`Most Memorable Characteristics`) %>%
  summarize(count=n()) %>%
  arrange(desc(count)) %>%
    wordcloud2(shape="circle") %>%
    htmlwidgets::prependContent(htmltools::tags$h2("Most Memorable Characteristics of 4-Rated Chocolate Bars"))
```


